The Mass Distribution of Stellar-Mass Black Holes

### 2010-11-05 paper.tex
As submitted.
higher-mass --> high-mass

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements, taking into account the number of degrees of freedom
  of each model.  The mass distribution of the low-mass systems is
  best fit by a power-law, while the distribution of the combined
  sample is best fit by the exponential model.  This difference
  indicates that the low-mass subsample is not consistent with being
  drawn from the distribution of the combined population.  We examine
  the existence of a ``gap'' between the most massive neutron stars
  and the least massive black holes by considering the value,
  $M_{1\%}$, of the 1\% quantile from each black hole mass
  distribution as the lower bound of black hole masses.  Our analysis
  generates posterior distributions for $M_{1\%}$; the best model (the
  power law) fitted to the low-mass systems has a distribution of
  lower-bounds with $M_{1\%} > 4.3$ $\Msun$ with 90\% confidence,
  while the best model (the exponential) fitted to all 20 systems has
  $M_{1\%} > 4.5$ $\Msun$ with 90\% confidence.  We conclude that our
  sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and the lower bound on black
  hole masses.  Our results on the low-mass sample are in qualitative
  agreement with those of \citet{Ozel2010}, although our broad
  model-selection analysis more reliably reveals the best-fit
  quantitative description of the underlying mass distribution.  The
  results on the combined sample of low- and higher-mass systems are
  in qualitative agreement with \citet{Fryer2001} although the
  presence of a mass gap remains theoretically unexplained.
\end{abstract}

@@

Observations of X-ray binaries in both the optical and X-ray bands can
provide a measurement of the mass of the compact object in these
systems.  The current sample of stellar mass black holes with
dynamically measured masses includes 15 systems with low-mass, Roche
lobe overflowing donors and 5 wind-fed systems with higher-mass
donors.  Hence, sophisticated statistical analyses of the black hole
mass distribution in these systems are possible.

@@

The structure of this paper is as follows.  In Section
\ref{sec:systems} we discuss the 15 systems that comprise the low-mass
X-ray binary black hole sample and the 5 additional higher-mass,
wind-fed systems that make up the combined sample.  In Section
\ref{sec:models} we discuss the Bayesian techniques we use to analyze
the black hole mass distribution, the techniques we use for model
selection, and the parametric and non-parametric models we will for
the black hole mass distribution.  In Section \ref{sec:results} we
discuss the results of our analysis and model selection.  In Section
\ref{sec:minimum-mass} we discuss the distribution of the minimum
black hole mass implied by the analysis of Section \ref{sec:results}.
In Section \ref{sec:conclusion} we summarize our results and comment
on the significance of the observed mass gap in the context of
theoretical models.  Appendix \ref{sec:mcmc} describes MCMC techniques
in some detail.  Appendix \ref{sec:reversible-jump-mcmc} explains our
novel algorithm for efficiently performing the reversible jump MCMC
computations used in the model comparisons of Section
\ref{sec:results} (but see also \citet{Farr2010}).

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 higher-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears to be more massive than the black hole).  The
low- and higher-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Section \ref{sec:results-low-mass}), and then the
combined sample of 20 systems (Section \ref{sec:higher-mass}).

@@

\begin{figure}
  \begin{center}
    \plotone{plots/all-masses} 
  \end{center}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/high-masses}
  \end{center}
  \caption{\label{fig:high-masses} Mass distributions for the
    wind-fed, higher-mass systems computed from the distributions on
    observed data in Table \ref{tab:sources} using Equation
    \eqref{eq:mass-function}.  (Similar to Figure
    \ref{fig:low-masses}.)  The asymmetry and long tails in these
    distributions are the result of the non-linear relationship
    between $M$, $f$, $q$, and $i$.}
\end{figure}

@@

Though we do not expect to find a second peak in the low-mass
distribution, we may find evidence of one when exploring the combined
low- and higher-mass samples.  To look for a second peak in the
black-hole mass distribution, we use a two-Gaussian model:
\begin{multline}
  \label{eq:two-gaussian-def}
  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
\end{multline}
The probability is a linear combination of two Gaussians with weights
$\alpha$ and $1-\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose
combined conditions on $\mu_i$ and $\sigma_i$ that ensure that most of
the mass probability lies below $40 \Msun$ with the prior
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
    \mu_2, \quad 0 \leq \alpha \leq 1 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
in Equation \eqref{eq:gaussian-prior-def}.

@@

In this section we discuss the results of our MCMC analysis of the
posterior distributions of parameters for the models in Sections
\ref{sec:parametric-models} and \ref{sec:non-parametric-models}.  We
also discuss model selection results.  The results in Section
\ref{sec:results-low-mass} apply to the low-mass sample of
systems, while those of Section \ref{sec:higher-mass} apply to the
combined sample of systems.

@@

\begin{figure}
  \begin{center}
    \plotone{plots/dist}
  \end{center}
  \caption{\label{fig:dists} The median (solid line), 10\% (lower
    dashed line) and 90\% (upper dashed line) values of the black hole
    mass distribution, $p(M|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{sec:parametric-models} and \ref{sec:non-parametric-models}.
    These distributions use only the 15 low-mass observations in Table
    \ref{tab:sources} (the combined sample is analyzed in Section
    \ref{sec:higher-mass}).  Note that these ``distributions of
    distributions'' are not necessarily normalized, and need not be
    shaped like the underlying model distributions.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/power-law}
  \end{center}
  \caption{\label{fig:power-law} Histograms of the marginalized
    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
    (top, right), and $\alpha$ (bottom) from the power-law model.  The
    marginalized distribution for $\alpha$ is broad, with $-11.8 <
    \alpha < 6.8$ enclosing 90\% of the probability.  We have
    $p(\alpha < 0) = 0.6$; the median value is $\alpha = -3.35$.  The
    broad distribution for $\alpha$ (and the other parameters) is due
    to correlations between the parameters discussed in the main text;
    see Figure \ref{fig:power-law-2D}.}
\end{figure}

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow range in mass of the
low-mass systems.  When $\alpha$ is negative, the resulting mass
distribution slopes down; $\Mmin$ is constrained to be near the lowest
mass of the observed black holes, while $\Mmax$ is essentially
irrelevant.  Conversely, when $\alpha$ is positive and the mass
distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the higher-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

\begin{figure}
  \begin{center}
    \plotone{plots/power-law-2D}
  \end{center}
  \caption{\label{fig:power-law-2D} MCMC samples in the $\Mmin,
    \alpha$ (top) and $\Mmax, \alpha$ (bottom) planes for the
    power-law model discussed in Section \ref{sec:power-law}.  The
    correlations between $\alpha$ and the power law bounds discussed
    in the text are apparent: when $\alpha$ is positive, the mass
    distribution slopes upward and $\Mmax$ is constrained to be near
    the maximum observed mass while $\Mmin$ is unconstrained.  When
    $\alpha$ is negative, the mass distribution slopes down and
    $\Mmin$ is constrained to be near the lowest mass observed, while
    $\Mmax$ is unconstrained. }
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/exp-cutoff}
  \end{center}
  \caption{\label{fig:exp-marginal} The distribution of scale masses,
    $M_0$ (dashed histogram), and minimum masses, $\Mmin$ (solid
    histogram), both measured in units of a solar mass for the
    exponential underlying mass distribution defined in Equation
    \eqref{eq:exp-def}.  The median scale mass is $M_0 = 1.55$, and
    $0.78 \leq M_0 \leq 3.25$ with 90\% confidence.}
\end{figure}

\begin{figure}
  \begin{center}
    \plotone{plots/exp-cutoff-2d}
  \end{center}
  \caption{\label{fig:exp-2D} The MCMC samples in the $\Mmin$, $M_0$
    plane for the decaying exponential underlying mass distribution
    model.  The slight correlation between smaller $\Mmin$ and larger
    $M_0$ is driven by the need to widen the mass distribution to
    encompass the peak of the measurements in Figure
    \ref{fig:low-masses} when the minimum mass decreases.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/gaussian}
  \end{center}
  \caption{\label{fig:gaussian} Marginalized posterior distributions
    for the mean, $\mu$ (solid histogram), and standard deviation,
    $\sigma$ (dashed histogram), both in solar masses for the Gaussian
    underlying mass distribution defined in Equation
    \eqref{eq:gaussian-def}.  The peak of the Gaussian, $\mu$, is
    constrained in $6.63 \leq \mu \leq 8.08$ with 90\% confidence.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  The top panel is
    $\mu_1$ (solid histogram) and $\sigma_1$ (dashed histogram), the
    middle panel is $\mu_2$ (solid histogram) and $\sigma_2$ (dashed
    histogram), and the bottom panel is $\alpha$. We have $\alpha >
    0.8$ with 62\% probability, favoring the first of the two
    Gaussians.  The distributions for $\mu_1$ and $\sigma_1$ are
    similar to those of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples.
    The extra degrees of freedom in the distribution from the second
    Gaussian do not provide enough extra fitting power to compensate
    for the increase in parmeter space, however: the two-Gaussian
    model is disfavored relative to the single Gaussian by a factor of
    $4.7$ on this dataset (see Sections \ref{sec:model-selection} and
    \ref{sec:low-mass-model-selection} for discussion).}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/log-normal}
  \end{center}
  \caption{\label{fig:log-normal} Marginalized distributions of the
    mean mass, $\langle M \rangle$ (solid histogram), and standard
    deviation of the mass, $\sigma_M$ (dashed histogram), for the
    log-normal model in Section \ref{sec:log-normal}.  The
    distributions are similar to the distributions of $\mu$ and
    $\sigma$ in the Gaussian model of Section \ref{sec:gaussian}.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/rj}
  \end{center}
  \caption{\label{fig:rj} The relative probability of the models
    discussed in Section \ref{sec:models} as computed using the
    reversible-jump MCMC with the efficient jump proposal algorithm
    described in Section \ref{sec:reversible-jump-mcmc}.  (See also
    Table \ref{tab:rj}.)  In increasing order along the $x$-axis, the
    models are the power-law of Section \ref{sec:power-law} (PL), the
    decaying exponential of Section \ref{sec:exponential} (E), the
    single Gaussian of Section \ref{sec:gaussian} (G), the double
    Gaussian of Section \ref{sec:gaussian} (TG), and the one-, two-,
    three-, four-, and five-bin histogram models of Section
    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
    respectively).  The average of 500 independent reversible-jump
    MCMCs is plotted, along with the 1-$\sigma$ error on the average
    inferred from the standard deviation of the probability from the
    individual MCMCs.  As discussed in the text, the power-law and
    Gaussian models are the most favored.}
\end{figure}

@@

\subsection{Combined Sample}
\label{sec:higher-mass}

This section repeats the analysis of the models from Section
\ref{sec:models}, but including the higher-mass, wind-fed systems from
Table \ref{tab:sources} (see also Figure \ref{fig:high-masses}) in the
sample.  Figure \ref{fig:high-mass-dists} displays bounds on the value
of the underlying mass distribution for the various models in Section
\ref{sec:models} applied to this data set; compare to Figure
\ref{fig:dists}.  The inclusion of the higher-mass, wind-fed systems
tends to widen the distribution toward the high-mass end and, in
models that allow it, produce a second, higher-mass peak in addition
to the one in Figure \ref{fig:dists}.

\begin{figure}
  \begin{center}
    \plotone{plots/dist-high}
  \end{center}
  \caption{\label{fig:high-mass-dists} The median (solid line), 10\%
    (lower dashed line), and 90\% (upper dashed line) values of the
    black hole mass distribution, $p(M|\theta)$, at various masses
    implied by the posterior $p(\theta|d)$ for the models discussed in
    Sections \ref{sec:parametric-models} and
    \ref{sec:non-parametric-models}.  These distributions use the
    combined sample of 20 observations in Table \ref{tab:sources},
    including the higher-mass, wind-fed systems.  Note that these
    ``distributions of distributions'' are not necessarily normalized,
    and need not be ``shaped'' like the underlying model
    distributions.  Compare to Figure \ref{fig:dists}, which includes
    only the low-mass systems in the analysis.  Including the
    higher-mass systems tends to widen the distribution toward the
    high-mass end and, in models that allow it, produce a second,
    higher-mass peak in addition to the one in Figure
    \ref{fig:dists}. }
\end{figure}

@@

Figure \ref{fig:power-law-high} presents the marginalized
distribution for the three power-law parameters $\Mmin$, $\Mmax$, and
$\alpha$ (Section \ref{sec:power-law}) from an analysis including the
higher-mass systems.  The distribution for $\Mmax$ is quite broad
because the best fit power laws slope downward ($\alpha < 0$), making
this parameter less relevant.  The range $-5.05 \leq \alpha \leq
-1.77$ encloses 90\% of the probability; the median value of $\alpha$
is -3.23.  The presence of the higher-mass samples in the analysis
produces a distinctive tail, eliminating the correlations discussed in
Section \ref{sec:power-law} and displayed in Figure
\ref{fig:power-law-2D} for the low-mass subset of the observations.

\begin{figure}
  \begin{center}
    \plotone{plots/power-law-high}
  \end{center}
  \caption{\label{fig:power-law-high} Histograms of the marginalized
    distribution for the three parameters $\Mmin$ (top, dashed
    histogram), $\Mmax$ (top, solid histogram), and $\alpha$ (bottom)
    from the power-law model including the higher-mass samples in the
    MCMC.  The distribution for $\Mmax$ is quite broad because the
    best fit power laws slope downward ($\alpha < 0$), making this
    parameter less relevant.  The range $-5.05 \leq \alpha \leq -1.77$
    encloses 90\% of the probability; the median value of $\alpha$ is
    -3.23.  The presence of the higher-mass samples in the analysis
    produces a distinctive tail, eliminating the correlations
    discussed in Section \ref{sec:power-law} and displayed in Figure
    \ref{fig:power-law-2D} for the low-mass subset of the
    observations. }
\end{figure}

@@

Figure \ref{fig:exp-cutoff-high} displays the marginalized
distributions for the exponential parameters $\Mmin$ and $M_0$
(Section \ref{sec:exponential}) from an analysis including the
higher-mass systems.  The distribution for the scale mass, $M_0$, has
moved to higher masses relative to Figure \ref{fig:exp-marginal} to
fit the tail of the mass distribution; the distribution for $\Mmin$ is
less affected, though it has broadened somewhat toward low masses.

>>>>

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  high-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements, taking into account the number of degrees of freedom
  of each model.  The mass distribution of the low-mass systems is
  best fit by a power-law, while the distribution of the combined
  sample is best fit by the exponential model.  This difference
  indicates that the low-mass subsample is not consistent with being
  drawn from the distribution of the combined population.  We examine
  the existence of a ``gap'' between the most massive neutron stars
  and the least massive black holes by considering the value,
  $M_{1\%}$, of the 1\% quantile from each black hole mass
  distribution as the lower bound of black hole masses.  Our analysis
  generates posterior distributions for $M_{1\%}$; the best model (the
  power law) fitted to the low-mass systems has a distribution of
  lower-bounds with $M_{1\%} > 4.3$ $\Msun$ with 90\% confidence,
  while the best model (the exponential) fitted to all 20 systems has
  $M_{1\%} > 4.5$ $\Msun$ with 90\% confidence.  We conclude that our
  sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and the lower bound on black
  hole masses.  Our results on the low-mass sample are in qualitative
  agreement with those of \citet{Ozel2010}, although our broad
  model-selection analysis more reliably reveals the best-fit
  quantitative description of the underlying mass distribution.  The
  results on the combined sample of low- and high-mass systems are
  in qualitative agreement with \citet{Fryer2001} although the
  presence of a mass gap remains theoretically unexplained.
\end{abstract}

@@

Observations of X-ray binaries in both the optical and X-ray bands can
provide a measurement of the mass of the compact object in these
systems.  The current sample of stellar mass black holes with
dynamically measured masses includes 15 systems with low-mass, Roche
lobe overflowing donors and 5 wind-fed systems with high-mass
donors.  Hence, sophisticated statistical analyses of the black hole
mass distribution in these systems are possible.

@@

The structure of this paper is as follows.  In Section
\ref{sec:systems} we discuss the 15 systems that comprise the low-mass
X-ray binary black hole sample and the 5 additional high-mass,
wind-fed systems that make up the combined sample.  In Section
\ref{sec:models} we discuss the Bayesian techniques we use to analyze
the black hole mass distribution, the techniques we use for model
selection, and the parametric and non-parametric models we will for
the black hole mass distribution.  In Section \ref{sec:results} we
discuss the results of our analysis and model selection.  In Section
\ref{sec:minimum-mass} we discuss the distribution of the minimum
black hole mass implied by the analysis of Section \ref{sec:results}.
In Section \ref{sec:conclusion} we summarize our results and comment
on the significance of the observed mass gap in the context of
theoretical models.  Appendix \ref{sec:mcmc} describes MCMC techniques
in some detail.  Appendix \ref{sec:reversible-jump-mcmc} explains our
novel algorithm for efficiently performing the reversible jump MCMC
computations used in the model comparisons of Section
\ref{sec:results} (but see also \citet{Farr2010}).

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 high-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears to be more massive than the black hole).  The
low- and high-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Section \ref{sec:results-low-mass}), and then the
combined sample of 20 systems (Section \ref{sec:high-mass}).

@@

\begin{figure}
  \begin{center}
    \plotone{all-masses} 
  \end{center}

@@

\begin{figure}
  \begin{center}
    \plotone{high-masses}
  \end{center}
  \caption{\label{fig:high-masses} Mass distributions for the
    wind-fed, high-mass systems computed from the distributions on
    observed data in Table \ref{tab:sources} using Equation
    \eqref{eq:mass-function}.  (Similar to Figure
    \ref{fig:low-masses}.)  The asymmetry and long tails in these
    distributions are the result of the non-linear relationship
    between $M$, $f$, $q$, and $i$.}
\end{figure}

@@

Though we do not expect to find a second peak in the low-mass
distribution, we may find evidence of one when exploring the combined
low- and high-mass samples.  To look for a second peak in the
black-hole mass distribution, we use a two-Gaussian model:
\begin{multline}
  \label{eq:two-gaussian-def}
  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
\end{multline}
The probability is a linear combination of two Gaussians with weights
$\alpha$ and $1-\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose
combined conditions on $\mu_i$ and $\sigma_i$ that ensure that most of
the mass probability lies below $40 \Msun$ with the prior
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
    \mu_2, \quad 0 \leq \alpha \leq 1 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
in Equation \eqref{eq:gaussian-prior-def}.

@@

In this section we discuss the results of our MCMC analysis of the
posterior distributions of parameters for the models in Sections
\ref{sec:parametric-models} and \ref{sec:non-parametric-models}.  We
also discuss model selection results.  The results in Section
\ref{sec:results-low-mass} apply to the low-mass sample of
systems, while those of Section \ref{sec:high-mass} apply to the
combined sample of systems.

@@

\begin{figure}
  \begin{center}
    \plotone{dist}
  \end{center}
  \caption{\label{fig:dists} The median (solid line), 10\% (lower
    dashed line) and 90\% (upper dashed line) values of the black hole
    mass distribution, $p(M|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{sec:parametric-models} and \ref{sec:non-parametric-models}.
    These distributions use only the 15 low-mass observations in Table
    \ref{tab:sources} (the combined sample is analyzed in Section
    \ref{sec:high-mass}).  Note that these ``distributions of
    distributions'' are not necessarily normalized, and need not be
    shaped like the underlying model distributions.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{power-law}
  \end{center}
  \caption{\label{fig:power-law} Histograms of the marginalized
    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
    (top, right), and $\alpha$ (bottom) from the power-law model.  The
    marginalized distribution for $\alpha$ is broad, with $-11.8 <
    \alpha < 6.8$ enclosing 90\% of the probability.  We have
    $p(\alpha < 0) = 0.6$; the median value is $\alpha = -3.35$.  The
    broad distribution for $\alpha$ (and the other parameters) is due
    to correlations between the parameters discussed in the main text;
    see Figure \ref{fig:power-law-2D}.}
\end{figure}

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow range in mass of the
low-mass systems.  When $\alpha$ is negative, the resulting mass
distribution slopes down; $\Mmin$ is constrained to be near the lowest
mass of the observed black holes, while $\Mmax$ is essentially
irrelevant.  Conversely, when $\alpha$ is positive and the mass
distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the high-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

\begin{figure}
  \begin{center}
    \plotone{power-law-2D}
  \end{center}
  \caption{\label{fig:power-law-2D} MCMC samples in the $\Mmin,
    \alpha$ (top) and $\Mmax, \alpha$ (bottom) planes for the
    power-law model discussed in Section \ref{sec:power-law}.  The
    correlations between $\alpha$ and the power law bounds discussed
    in the text are apparent: when $\alpha$ is positive, the mass
    distribution slopes upward and $\Mmax$ is constrained to be near
    the maximum observed mass while $\Mmin$ is unconstrained.  When
    $\alpha$ is negative, the mass distribution slopes down and
    $\Mmin$ is constrained to be near the lowest mass observed, while
    $\Mmax$ is unconstrained. }
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{exp-cutoff}
  \end{center}
  \caption{\label{fig:exp-marginal} The distribution of scale masses,
    $M_0$ (dashed histogram), and minimum masses, $\Mmin$ (solid
    histogram), both measured in units of a solar mass for the
    exponential underlying mass distribution defined in Equation
    \eqref{eq:exp-def}.  The median scale mass is $M_0 = 1.55$, and
    $0.78 \leq M_0 \leq 3.25$ with 90\% confidence.}
\end{figure}

\begin{figure}
  \begin{center}
    \plotone{exp-cutoff-2d}
  \end{center}
  \caption{\label{fig:exp-2D} The MCMC samples in the $\Mmin$, $M_0$
    plane for the decaying exponential underlying mass distribution
    model.  The slight correlation between smaller $\Mmin$ and larger
    $M_0$ is driven by the need to widen the mass distribution to
    encompass the peak of the measurements in Figure
    \ref{fig:low-masses} when the minimum mass decreases.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{gaussian}
  \end{center}
  \caption{\label{fig:gaussian} Marginalized posterior distributions
    for the mean, $\mu$ (solid histogram), and standard deviation,
    $\sigma$ (dashed histogram), both in solar masses for the Gaussian
    underlying mass distribution defined in Equation
    \eqref{eq:gaussian-def}.  The peak of the Gaussian, $\mu$, is
    constrained in $6.63 \leq \mu \leq 8.08$ with 90\% confidence.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  The top panel is
    $\mu_1$ (solid histogram) and $\sigma_1$ (dashed histogram), the
    middle panel is $\mu_2$ (solid histogram) and $\sigma_2$ (dashed
    histogram), and the bottom panel is $\alpha$. We have $\alpha >
    0.8$ with 62\% probability, favoring the first of the two
    Gaussians.  The distributions for $\mu_1$ and $\sigma_1$ are
    similar to those of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples.
    The extra degrees of freedom in the distribution from the second
    Gaussian do not provide enough extra fitting power to compensate
    for the increase in parmeter space, however: the two-Gaussian
    model is disfavored relative to the single Gaussian by a factor of
    $4.7$ on this dataset (see Sections \ref{sec:model-selection} and
    \ref{sec:low-mass-model-selection} for discussion).}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{log-normal}
  \end{center}
  \caption{\label{fig:log-normal} Marginalized distributions of the
    mean mass, $\langle M \rangle$ (solid histogram), and standard
    deviation of the mass, $\sigma_M$ (dashed histogram), for the
    log-normal model in Section \ref{sec:log-normal}.  The
    distributions are similar to the distributions of $\mu$ and
    $\sigma$ in the Gaussian model of Section \ref{sec:gaussian}.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{rj}
  \end{center}
  \caption{\label{fig:rj} The relative probability of the models
    discussed in Section \ref{sec:models} as computed using the
    reversible-jump MCMC with the efficient jump proposal algorithm
    described in Section \ref{sec:reversible-jump-mcmc}.  (See also
    Table \ref{tab:rj}.)  In increasing order along the $x$-axis, the
    models are the power-law of Section \ref{sec:power-law} (PL), the
    decaying exponential of Section \ref{sec:exponential} (E), the
    single Gaussian of Section \ref{sec:gaussian} (G), the double
    Gaussian of Section \ref{sec:gaussian} (TG), and the one-, two-,
    three-, four-, and five-bin histogram models of Section
    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
    respectively).  The average of 500 independent reversible-jump
    MCMCs is plotted, along with the 1-$\sigma$ error on the average
    inferred from the standard deviation of the probability from the
    individual MCMCs.  As discussed in the text, the power-law and
    Gaussian models are the most favored.}
\end{figure}

@@

\subsection{Combined Sample}
\label{sec:high-mass}

This section repeats the analysis of the models from Section
\ref{sec:models}, but including the high-mass, wind-fed systems from
Table \ref{tab:sources} (see also Figure \ref{fig:high-masses}) in the
sample.  Figure \ref{fig:high-mass-dists} displays bounds on the value
of the underlying mass distribution for the various models in Section
\ref{sec:models} applied to this data set; compare to Figure
\ref{fig:dists}.  The inclusion of the high-mass, wind-fed systems
tends to widen the distribution toward the high-mass end and, in
models that allow it, produce a second, high-mass peak in addition
to the one in Figure \ref{fig:dists}.

\begin{figure}
  \begin{center}
    \plotone{dist-high}
  \end{center}
  \caption{\label{fig:high-mass-dists} The median (solid line), 10\%
    (lower dashed line), and 90\% (upper dashed line) values of the
    black hole mass distribution, $p(M|\theta)$, at various masses
    implied by the posterior $p(\theta|d)$ for the models discussed in
    Sections \ref{sec:parametric-models} and
    \ref{sec:non-parametric-models}.  These distributions use the
    combined sample of 20 observations in Table \ref{tab:sources},
    including the high-mass, wind-fed systems.  Note that these
    ``distributions of distributions'' are not necessarily normalized,
    and need not be ``shaped'' like the underlying model
    distributions.  Compare to Figure \ref{fig:dists}, which includes
    only the low-mass systems in the analysis.  Including the
    high-mass systems tends to widen the distribution toward the
    high-mass end and, in models that allow it, produce a second,
    high-mass peak in addition to the one in Figure
    \ref{fig:dists}. }
\end{figure}

@@

Figure \ref{fig:power-law-high} presents the marginalized
distribution for the three power-law parameters $\Mmin$, $\Mmax$, and
$\alpha$ (Section \ref{sec:power-law}) from an analysis including the
high-mass systems.  The distribution for $\Mmax$ is quite broad
because the best fit power laws slope downward ($\alpha < 0$), making
this parameter less relevant.  The range $-5.05 \leq \alpha \leq
-1.77$ encloses 90\% of the probability; the median value of $\alpha$
is -3.23.  The presence of the high-mass samples in the analysis
produces a distinctive tail, eliminating the correlations discussed in
Section \ref{sec:power-law} and displayed in Figure
\ref{fig:power-law-2D} for the low-mass subset of the observations.

\begin{figure}
  \begin{center}
    \plotone{power-law-high}
  \end{center}
  \caption{\label{fig:power-law-high} Histograms of the marginalized
    distribution for the three parameters $\Mmin$ (top, dashed
    histogram), $\Mmax$ (top, solid histogram), and $\alpha$ (bottom)
    from the power-law model including the high-mass samples in the
    MCMC.  The distribution for $\Mmax$ is quite broad because the
    best fit power laws slope downward ($\alpha < 0$), making this
    parameter less relevant.  The range $-5.05 \leq \alpha \leq -1.77$
    encloses 90\% of the probability; the median value of $\alpha$ is
    -3.23.  The presence of the high-mass samples in the analysis
    produces a distinctive tail, eliminating the correlations
    discussed in Section \ref{sec:power-law} and displayed in Figure
    \ref{fig:power-law-2D} for the low-mass subset of the
    observations. }
\end{figure}

@@

Figure \ref{fig:exp-cutoff-high} displays the marginalized
distributions for the exponential parameters $\Mmin$ and $M_0$
(Section \ref{sec:exponential}) from an analysis including the
high-mass systems.  The distribution for the scale mass, $M_0$, has
moved to higher masses relative to Figure \ref{fig:exp-marginal} to
fit the tail of the mass distribution; the distribution for $\Mmin$ is
less affected, though it has broadened somewhat toward low masses.
